Solar wind predictions for the Parker Solar Probe orbit

### 2017-11-10 paperMVVB_for_aanda/paperMVVB.tex
cleared tex files from comments; integrated all into one tex file

\section{Introduction}
%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1959 \citep{Gringauz1960,Neugebauer1966}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{140}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} in October 2017, having crossed the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
Various spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with WIND \citep{Lepping1995,Ogilvie1995}, SOHO \citep{Domingo1995}, ACE \citep{Stone1998} and DSCOVR \citep{Burt2012} currently providing observations while orbiting around the L1 point \SI{1.5}{million\,\km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \citep{Colin1980} or MESSENGER \citep{Belcher1991}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. The in situ solar wind measurements closest to the Sun to date were made by the Helios missions. Helios~1, launched in 1974, reached distances of \SI{0.31}{\au}. Helios~2, launched two years later, approached the Sun as close as \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). In comparison, the ESA Solar Orbiter mission with a planned launch in February 2019 will have its closest perihelia at \SI{0.28}{\au} \citep{Muller2013}.

%PSP objectives and instruments
The key PSP science objectives are to “trace the flow of energy that heats and accelerates the solar corona and solar wind, determine the structure and dynamics of the plasma and magnetic fields at the sources of the solar wind, and explore mechanisms that accelerate and transport energetic particles” as stated in \citet{Fox2015}. To achieve these goals, PSP has four scientific instruments on board: FIELDS for the measurements of magnetic fields and AC/DC electric fields \citep{Bale2016}, SWEAP for the measurements of flux of electrons, protons and alphas \citep{Kasper2016}, IS\sun{}IS for the measurement of solar energetic particles (SEPs) \citep{McComas2016} and WISPR for the measurement of coronal and inner heliospheric structures \citep{Vourlidas2016}.

%CGAUSS and WISPR
The study presented in this paper is undertaken in the Coronagraphic German And US SolarProbePlus Survey (CGAUSS) project, which is the German contribution to the PSP mission as part of the Wide-field Imager for Solar PRobe (WISPR). WISPR will contribute to the PSP science goals by deriving the 3D structure of the solar corona through which the in situ measurements are made to determine the sources of the solar wind. It will provide density power spectra over a wide range of structures (e.g., streamers, pseudostreamers and equatorial coronal holes) for determining the roles of turbulence, waves and pressure-balanced structures in the solar wind. It will also measure the physical properties, such as speed and density jumps of SEP-producing shocks and their coronal mass ejection (CME) drivers as they evolve in the corona and inner heliosphere \citep{Vourlidas2016}.

@@

%definition of sw environment
%section{Solar wind}
Generally two types of solar wind are observed in the heliosphere -- slow and fast streams \citep{Neugebauer1966,Schwenn1983}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses' out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types \citep{Schwenn1983}, is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, such as intermittent reconnection at the top of helmet streamers and from coronal hole boundaries \citep{Kilpua2016}. The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions which may persist for many solar rotations ("co-rotating" interaction regions) if the coronal source regions are quasi-stationary \citep{Balogh1999}.
Embedded in the slow and fast solar wind streams are transient flows of CMEs -- the faster ones driving shock waves ahead \citep{Gosling1974}. Their rate follows the solar activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

%solar wind parameters
It is not known which specific solar wind type or structure PSP will encounter at a given time during its mission, therefore we extrapolate the probability distributions of the major solar wind parameters from existing solar wind measurements and take solar cycle dependencies into account.
As a baseline we describe the solar wind environment through the key quantities of a magnetized plasma: \emph{magnetic field strength}, \emph{density} and \emph{temperature}. Furthermore, the bulk flow \emph{velocity} is the defining parameter of the two types of solar wind. Solar wind quantities, like flux densities, mass flux and plasma beta, can directly be derived from these four parameters. In the analyses we treat the solar wind as a proton plasma -- the average helium abundance is about \SI{4.5}{\percent} and in slow wind at solar cycle minimum even less than \SI{2}{\percent} \citep{Feldman1978,Schwenn1983,Kasper2012}.

%sections
Our approach is to obtain analytical representations of the shapes of the solar wind parameter’s frequency distributions in Sect.~\ref{sec:frequency_distribution}, of their solar activity dependence in Sect.~\ref{sec:solar_activity_variations} and of their solar distance scaling in Sect.~\ref{sec:solar_distance_dependency}. The solar wind parameters’ frequency distributions and solar activity dependence is derived from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles. Their distance dependency is derived from Helios solar wind measurements covering more than two third of the distance to the Sun and more than half a solar cycle. From combination of the obtained frequency distributions, SSN dependence functions and solar distance dependence functions a general solar wind model is build in Sect.~\ref{sec:empirical_solar_wind_model}, representing the solar activity and distance behavior. Finally, this empirical model is fed with SSN predictions and extrapolated to PSP's planned orbital positions in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}.

%lognormal fitting
\section{Frequency distributions of the solar wind parameters}
\label{sec:frequency_distribution}
The solar wind parameters are highly variable due to short-term variations from structures like slow and fast wind streams, interaction regions and CMEs, whose rate and properties depend on the phase of the solar activity cycle. Hence, for deriving characteristic frequency distributions for the solar wind parameters, measurements over long-term time spans are needed. The abundance of the near-Earth hourly OMNI data set is ideally suited for this purpose, because it spans to date almost five solar cycles.

%about OMNI data
The OMNI~2 data set \citep{King2005} combines solar wind magnetic field and plasma data collected by various satellites since 1963, currently by WIND and by ACE. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the different parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\
%
%hourly data
We note that a test-comparison of hourly averaged with higher time resolution data for the available shorter time span 1981--2016 did not show significant differences in our results.

%data binning
According to the OMNI data precision and maximal parameter ranges we specify bin sizes of \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature. The frequency distributions of the solar wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
The solar wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, and the temperature in the range \SIrange{3450}{6.63e6}{\K}, the mean data values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).
Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 CME with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013} or the solar wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.

%subsection{Lognormal fitting}
%choice of distribution function

@@

% Vdbl fit:
% medians = 489, 368
% means = 500, 372
% c = 0.504
% via calculate_twolognormal_median.py:
% mean = 442.38
% median = 416.05
% fit data difference (normed area):\\
% B: 0.0682978\\
% v: 0.186874 -> 0.0419676 (double fit)\\
% n: 0.0648273\\
% T: 0.0578349\\

@@

%double lognormal fitting
%%%%%%%%%%%%%%%%%%%%%%%%%
%justification for compositional lognormal fit
In order to find a better fit result for the velocity distribution, we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
\begin{align}
	W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.	\label{eq:double_lognormal_fit_function}
\end{align}
The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.
The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0	\,.
\end{align}
This more complex fit function is more accurate in describing the velocity's frequency distribution as shown in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{Velocity frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset is a zoomed-in view of the high value tail of the distribution.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
Thus in the following sections we keep the double lognormal ansatz for all velocity frequency fits.

%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well. The abnormally high parameter values in the distribution functions can be attributed to shock/CME events in agreement with the results of the OMNI solar wind investigations by \citet{Richardson2012}. The simple lognormal fit functions underestimate the frequencies in their high value tails, except for the temperature’s tail which is overestimated, as seen in the insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}, this appears to be because CMEs do not come with abnormally high temperatures, but rather with temperatures lower than those of the average solar wind \citep{Forsyth2006}. The velocity's compositional lognormal fit only slightly overestimates its tail as seen in the inset of Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
The slow and fast part contribute almost equally ($c \approx 0.5$) to the long-term velocity distribution function.

%{\color{red} discuss high value zoom figures; read in Veselovsky2010}

@@

%\subsection{SSN correlation}
For the correlations we fit lognormal functions to the frequency distributions as in Sect.~\ref{sec:frequency_distribution}, but implement linear relations to the yearly SSN, allowing shifting of the distribution functions with SSN. For the velocity the approach is different insofar as its two components are kept fixed and instead their balance is modified with the changing SSN. Thus we obtain solar activity depending models for the frequency distributions of all four solar wind parameters.

%\subsection{SSN data}
The international sunspot number (\citeyear{sidc}) is provided by the online catalogue\footnote{\url{http://www.sidc.be/silso/}} at the World Data Center -- Sunspot Index and Long-term Solar Observations (WDC-SILSO), Solar Influences Data Analysis Center (SIDC), Royal Observatory of Belgium (ROB).

Yearly medians of the solar wind parameters and the yearly SSN together with the solar cycle number are shown in the upper part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}. The reason for correlating the SSN to the solar wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{Solar wind parameter yearly frequencies (grey shading) with yearly medians (lines) derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}
The solar wind velocity, density and temperature depend on the state of the solar cycle \citep{Schwenn1983}. %density modulation with SSN \citep{Schwenn1983} p.~499\\
For instance the fast solar wind occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle as pointed out by \citet[p.~75, Figure~3.52]{Bothmer2007}. Therefore the solar wind velocity, density and temperature maxima exhibit time lags relative to the SSN maxima.

>>>>

\documentclass[]{aa}

\usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{multirow}
\newcommand{\Rs}{$R_\sun{}$}
\usepackage{siunitx}
	\DeclareSIUnit[number-unit-product=\,]\au{au}
	\DeclareSIUnit[number-unit-product=\,]\nT{\nano\tesla}
	\DeclareSIUnit[number-unit-product=\,]\Rs{\textit{R}_\sun{}}
	\sisetup{table-figures-uncertainty=2, table-number-alignment=center, range-phrase=--, range-units=single}

\begin{document}
	\title{Solar wind predictions for the Parker Solar Probe orbit}
	\subtitle{Near-Sun extrapolations derived from an empirical solar wind model based on Helios and OMNI observations}
	\author{M.~S.~Venzmer
	\and V.~Bothmer}
	\institute{University of Goettingen, Institute for Astrophysics, Friedrich-Hund-Platz~1, 37077~Göttingen, Germany}
	\date{Received 25 August 2017; accepted date}

	\abstract
	{The Parker Solar Probe (PSP) (formerly Solar Probe Plus) mission will be humanity’s first in situ exploration of the solar corona with closest perihelia at \num{9.86}~solar radii (\si{\Rs}) distance to the Sun. It will help answer hitherto unresolved questions on the heating of the solar corona and the source and acceleration of the solar wind and solar energetic particles. The scope of this study is to model the solar wind environment for PSP’s unprecedented distances in its prime mission phase during the years \numrange{2018}{2025}. The study is performed within the project Coronagraphic German And US SolarProbePlus Survey (CGAUSS) which is the German contribution to the PSP mission as part of the Wide-field Imager for Solar PRobe (WISPR).}
	{We present an empirical solar wind model for the inner heliosphere which is derived from OMNI and Helios data. The German-US space probes Helios~1 and Helios~2 flew in the 1970s and observed solar wind in the ecliptic within heliocentric distances of \SIrange{0.29}{0.98}{\au}. The OMNI database consists of multi-spacecraft intercalibrated in situ data obtained near \SI{1}{\au} over more than five solar cycles. The international sunspot number (SSN) and its predictions are used to derive dependencies of the major solar wind parameters on solar activity and to forecast their properties for the PSP mission.}
	{The frequency distributions for the solar wind key parameters magnetic field strength, proton velocity, density and temperature are represented by lognormal functions. In addition, we consider the velocity distribution’s bi-componental shape, consisting of a slower and a faster part. Functional relations to solar activity are compiled with use of the OMNI data by correlating and fitting the frequency distributions with the SSN. Further, based on the combined data set from both Helios probes, the parameters’ frequency distributions are fitted with respect to solar distance to obtain power law dependencies. Thus an empirical solar wind model for the inner heliosphere confined to the ecliptic region is derived, accounting for solar activity and for solar distance through adequate shifts of the lognormal distributions. Finally, the inclusion of SSN predictions and the extrapolation down to PSP’s perihelion region enables us to estimate the solar wind environment for PSP’s planned trajectory during its mission duration.}
	{The CGAUSS empirical solar wind model for PSP yields dependencies on solar activity and solar distance for the solar wind parameters' frequency distributions. The estimated solar wind median values for PSP’s first perihelion in 2018 at a solar distance of \SI{0.16}{\au} are \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{214}{\per\cm\cubed} and \SI{503000}{\K}. The estimates for PSP’s first closest perihelion, occuring in 2024 at \SI{0.046}{\au} (\SI{9.86}{\Rs}), are \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{2951}{\per\cm\cubed} and \SI{1930000}{\K}. Since the modeled velocity and temperature values below about \SI{20}{\Rs} appear overestimated in comparison with existing observations, this suggests that PSP will directly measure solar wind acceleration and heating processes below \SI{20}{\Rs} as planned.}
	{}
	\keywords{solar wind -- sun: heliosphere -- sun: corona}

	\maketitle

	\section{Introduction}
	From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1959 \citep{Gringauz1960,Neugebauer1966}.
	The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{140}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} in October 2017, having crossed the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
	Various spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with WIND \citep{Lepping1995,Ogilvie1995}, SOHO \citep{Domingo1995}, ACE \citep{Stone1998} and DSCOVR \citep{Burt2012} currently providing observations while orbiting around the L1 point \SI{1.5}{million\,\km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \citep{Colin1980} or MESSENGER \citep{Belcher1991}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. The in situ solar wind measurements closest to the Sun to date were made by the Helios missions. Helios~1, launched in 1974, reached distances of \SI{0.31}{\au}. Helios~2, launched two years later, approached the Sun as close as \SI{0.29}{\au} \citep{Rosenbauer1977}.
	The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). In comparison, the ESA Solar Orbiter mission with a planned launch in February 2019 will have its closest perihelia at \SI{0.28}{\au} \citep{Muller2013}.

	The key PSP science objectives are to “trace the flow of energy that heats and accelerates the solar corona and solar wind, determine the structure and dynamics of the plasma and magnetic fields at the sources of the solar wind, and explore mechanisms that accelerate and transport energetic particles” as stated in \citet{Fox2015}. To achieve these goals, PSP has four scientific instruments on board: FIELDS for the measurements of magnetic fields and AC/DC electric fields \citep{Bale2016}, SWEAP for the measurements of flux of electrons, protons and alphas \citep{Kasper2016}, IS\sun{}IS for the measurement of solar energetic particles (SEPs) \citep{McComas2016} and WISPR for the measurement of coronal and inner heliospheric structures \citep{Vourlidas2016}.

	The study presented in this paper is undertaken in the Coronagraphic German And US SolarProbePlus Survey (CGAUSS) project, which is the German contribution to the PSP mission as part of the Wide-field Imager for Solar PRobe (WISPR). WISPR will contribute to the PSP science goals by deriving the 3D structure of the solar corona through which the in situ measurements are made to determine the sources of the solar wind. It will provide density power spectra over a wide range of structures (e.g., streamers, pseudostreamers and equatorial coronal holes) for determining the roles of turbulence, waves and pressure-balanced structures in the solar wind. It will also measure the physical properties, such as speed and density jumps of SEP-producing shocks and their coronal mass ejection (CME) drivers as they evolve in the corona and inner heliosphere \citep{Vourlidas2016}.

@@

	Generally two types of solar wind are observed in the heliosphere -- slow and fast streams \citep{Neugebauer1966,Schwenn1983}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses' out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types \citep{Schwenn1983}, is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, such as intermittent reconnection at the top of helmet streamers and from coronal hole boundaries \citep{Kilpua2016}. The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions which may persist for many solar rotations ("co-rotating" interaction regions) if the coronal source regions are quasi-stationary \citep{Balogh1999}.
	Embedded in the slow and fast solar wind streams are transient flows of CMEs -- the faster ones driving shock waves ahead \citep{Gosling1974}. Their rate follows the solar activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

	It is not known which specific solar wind type or structure PSP will encounter at a given time during its mission, therefore we extrapolate the probability distributions of the major solar wind parameters from existing solar wind measurements and take solar cycle dependencies into account.
	As a baseline we describe the solar wind environment through the key quantities of a magnetized plasma: \emph{magnetic field strength}, \emph{density} and \emph{temperature}. Furthermore, the bulk flow \emph{velocity} is the defining parameter of the two types of solar wind. Solar wind quantities, like flux densities, mass flux and plasma beta, can directly be derived from these four parameters. In the analyses we treat the solar wind as a proton plasma -- the average helium abundance is about \SI{4.5}{\percent} and in slow wind at solar cycle minimum even less than \SI{2}{\percent} \citep{Feldman1978,Schwenn1983,Kasper2012}.

	Our approach is to obtain analytical representations of the shapes of the solar wind parameter’s frequency distributions in Sect.~\ref{sec:frequency_distribution}, of their solar activity dependence in Sect.~\ref{sec:solar_activity_variations} and of their solar distance scaling in Sect.~\ref{sec:solar_distance_dependency}. The solar wind parameters’ frequency distributions and solar activity dependence is derived from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles. Their distance dependency is derived from Helios solar wind measurements covering more than two third of the distance to the Sun and more than half a solar cycle. From combination of the obtained frequency distributions, SSN dependence functions and solar distance dependence functions a general solar wind model is build in Sect.~\ref{sec:empirical_solar_wind_model}, representing the solar activity and distance behavior. Finally, this empirical model is fed with SSN predictions and extrapolated to PSP's planned orbital positions in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}.

	\section{Frequency distributions of the solar wind parameters}
	\label{sec:frequency_distribution}
	The solar wind parameters are highly variable due to short-term variations from structures like slow and fast wind streams, interaction regions and CMEs, whose rate and properties depend on the phase of the solar activity cycle. Hence, for deriving characteristic frequency distributions for the solar wind parameters, measurements over long-term time spans are needed. The abundance of the near-Earth hourly OMNI data set is ideally suited for this purpose, because it spans to date almost five solar cycles.

	The OMNI~2 data set \citep{King2005} combines solar wind magnetic field and plasma data collected by various satellites since 1963, currently by WIND and by ACE. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
	In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the different parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
	We note that a test-comparison of hourly averaged with higher time resolution data for the available shorter time span 1981--2016 did not show significant differences in our results.

	According to the OMNI data precision and maximal parameter ranges we specify bin sizes of \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature. The frequency distributions of the solar wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
	The solar wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, and the temperature in the range \SIrange{3450}{6.63e6}{\K}, the mean data values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).
	Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 CME with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013} or the solar wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.

@@

@@

	In order to find a better fit result for the velocity distribution, we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
	\begin{align}
		W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.	\label{eq:double_lognormal_fit_function}
	\end{align}
	The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.
	The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived via solving
	\begin{align}
		\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0	\,.
	\end{align}
	This more complex fit function is more accurate in describing the velocity's frequency distribution as shown in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
	\begin{figure}
		\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
		\caption{Velocity frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset is a zoomed-in view of the high value tail of the distribution.}
		\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
	\end{figure}
	Thus in the following sections we keep the double lognormal ansatz for all velocity frequency fits.

	For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well. The abnormally high parameter values in the distribution functions can be attributed to shock/CME events in agreement with the results of the OMNI solar wind investigations by \citet{Richardson2012}. The simple lognormal fit functions underestimate the frequencies in their high value tails, except for the temperature’s tail which is overestimated, as seen in the insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}, this appears to be because CMEs do not come with abnormally high temperatures, but rather with temperatures lower than those of the average solar wind \citep{Forsyth2006}. The velocity's compositional lognormal fit only slightly overestimates its tail as seen in the inset of Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
	The slow and fast part contribute almost equally ($c \approx 0.5$) to the long-term velocity distribution function.

@@

	For the correlations we fit lognormal functions to the frequency distributions as in Sect.~\ref{sec:frequency_distribution}, but implement linear relations to the yearly SSN, allowing shifting of the distribution functions with SSN. For the velocity the approach is different insofar as its two components are kept fixed and instead their balance is modified with the changing SSN. Thus we obtain solar activity depending models for the frequency distributions of all four solar wind parameters.

	The international sunspot number (\citeyear{sidc}) is provided by the online catalogue\footnote{\url{http://www.sidc.be/silso/}} at the World Data Center -- Sunspot Index and Long-term Solar Observations (WDC-SILSO), Solar Influences Data Analysis Center (SIDC), Royal Observatory of Belgium (ROB).

	Yearly medians of the solar wind parameters and the yearly SSN together with the solar cycle number are shown in the upper part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}. The reason for correlating the SSN to the solar wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
	\begin{figure}
		\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
		\caption{Solar wind parameter yearly frequencies (grey shading) with yearly medians (lines) derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
		\label{fig:OMNI_yearly_ssn_correlation_c_plot}
	\end{figure}
	The solar wind velocity, density and temperature depend on the state of the solar cycle \citep{Schwenn1983}.
	For instance the fast solar wind occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle as pointed out by \citet[p.~75, Figure~3.52]{Bothmer2007}. Therefore the solar wind velocity, density and temperature maxima exhibit time lags relative to the SSN maxima.
